---
title: "simpleRegress"
author: "Danitza Nunez"
date: "Fall 11/16/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## make a simple regression model

```{r}
bears <- read.csv(file="BEARS.csv",header = TRUE)
#adjust factors to be factors
bears$SEX<- as.factor(bears$SEX)
bears$MONTH<- as.factor(bears$MONTH)
```

## Model LENGTH ~ NECK  this is an example  

```{r}
simp <- lm(LENGTH~NECK,data=bears)
plot(LENGTH~NECK,data=bears)
abline(simp)
summary.lm(simp)
```

## now predict the length of a bear who has a neck of 17, 22 and 26


```{r}
new <- data.frame(NECK=c(17,22,26))
predict(simp,new)
predict(simp,new,se.fit=TRUE)
```
## Surely there is another variable that can better predict LENGTH than NECK. Find one and compare its performance to that of neck by it value of Adjusted-Rsquared (bigger is better).

```{r}
simp2 <- lm(LENGTH~HEADWTH,data=bears)
plot(LENGTH~HEADWTH,data=bears)
abline(simp2)
summary.lm(simp2)
```

The adjusted R-squared is 0.53 which is smaller than the model using "neck". 



```{r}
simp3 <- lm(LENGTH~WEIGHT,data=bears)
plot(LENGTH~WEIGHT,data=bears)
abline(simp3)
summary.lm(simp3)
```

The Adjusted R-squared is 0.7423 approximately the same as the modle "neck", hence not bigger. While attempt a different variable. 


```{r}
simp4 <- lm(LENGTH~CHEST,data=bears)
plot(LENGTH~CHEST,data=bears)
abline(simp4)
summary.lm(simp4)
```

The Adjusted R-squared of the chest variable is 0.7869 which is larger than the model "neck" variable. 


```{r}
simp5 <- lm(LENGTH~HEADLEN,data=bears)
plot(LENGTH~HEADLEN,data=bears)
abline(simp5)
summary.lm(simp5)
```
The Adjusted R-squared of HEADLEN is  0.8434 higher than the model "neck" variable. 


```{r}
simp6 <- lm(LENGTH~as.numeric(SEX),data=bears)
plot(LENGTH~as.numeric(SEX),data=bears)
abline(simp6)
summary.lm(simp6)
```
The Adjusted R-squared of the SEX variable  is -0.008373 which is too small to consider. 



```{r}
simp7 <- lm(LENGTH~as.numeric(MONTH),data=bears)
plot(LENGTH~as.numeric(MONTH),data=bears)
abline(simp7)
summary.lm(simp7)
```
the Adjusted R-squared for the MONTH variable is  -0.01153  compared to model "NECK" which is too small to consider. 